---
title: "(Many) Schools in RStan"
output: html_notebook
---

Takes the 8Schools example from the RStan [GitHub page][1] and allows for increasing the number of schools, testing capacity of a machine.

First load the `rstan` library, set some options and load the original data used in the example for 8 schools.

```{r}
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

schools_dat <- list(J = 8, 
                    y = c(28, 8, -3, 7, -1, 1, 18, 12), 
                    sigma = c(15, 10, 16, 11, 9, 11, 10, 18))
```

Now set parameter `big_school_size`, and use it to generate a larger dataset.

```{r}
big_school_size <- 50L

big_schools_dat <- list(
    J = big_school_size, 
    y = sample(schools_dat$y, size = big_school_size, replace = TRUE), 
    sigma = sample(schools_dat$sigma, size = big_school_size, replace = TRUE)
)
```

Now fit the `stan` model. This calls from the `8schools.stan` script saved locally. Parameters `iter` and `chains` will affect the time taken to run (controlling the number of iterations and [Markov Chains][2] respectively). Wrapping it in `system.time` captures the time taken. NB. that elapsed time will be more accurate if using multiple cores.

```{r}
system.time(
    {
        fit <- stan(file = "8schools.stan", 
                    data = big_schools_dat, 
                    iter = 10000, 
                    chains = 8, 
                    control = list(adapt_delta = 0.9))
    }
)
```

Several built-in methods available for printing and plotting.
```{r}
print(fit)
plot(fit)
pairs(fit, pars = c("mu", "tau", "lp__"))
```

Some examples of manipulations to get objects for use in other applications.

```{r}
la <- extract(fit, permuted = TRUE)
mu <- la$mu
a <- extract(fit, permuted = FALSE)
a2 <- as.array(fit)
m <- as.matrix(fit)
d <- as.data.frame(fit)

print(fit, digits = 1)
```

[1]: https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started
[2]: https://en.wikipedia.org/wiki/Markov_chain