---
title: "Testing R on Stride"
output: html_notebook
---
Notebook for testing R on STRIDE. Runs through downloading data, plotting various elements and then running machine learning algorithms to predict outcome.

Dataset used is the [UCI Adult][1] data, which allows for binary prediction of whether or not given adult has an income above $50,000.

```{r load libraries, message=FALSE, warning=FALSE, paged.print=FALSE}
library(purrr)
# library(data.table)
library(dplyr)
library(ggplot2)
# library(ggthemes)
library(tidyr)
library(readr)
library(caret)
library(xgboost)
library(Matrix)
```

```{r get data}
data_url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data"
data_filepath <- "./data/adult.csv"
if (!file.exists(data_filepath)) {
    download.file(data_url, data_filepath)
}
headers <- c("age", "workclass", "fnlwgt", "education", "education_num", 
             "marital_status", "occupation", "relationship", "race", 
             "sex", "cap_gain", "cap_loss", "hrs_per_week", "native_country", 
             "income_level")
df <- read_csv(data_filepath, col_names = headers)
```

Now do some quick plotting to check for obvious problems with the dataset (e.g. imbalanced classes).

```{r plotting variables}
df %>% 
    dplyr::select_if(is.character) %>% 
    gather() %>% 
    ggplot(aes(x = value)) +
    geom_bar() +
    facet_wrap(~ key, scales = "free_x") + 
    theme_minimal() +
    # theme_tufte() +
    theme(axis.text.x = element_blank())

df %>% 
    dplyr::select_if(is.integer) %>% 
    gather() %>% 
    ggplot(aes(x = value)) +
    geom_histogram(bins = 20) +
    facet_wrap(~ key, scales = "free_x") + 
    theme_minimal() +
    # theme_tufte() +
    theme(axis.text.x = element_blank())
```

Check for, and handle, missing data.

```{r missing data}
na_counts <- sapply(df, function(y) {
    sum(length(which(is.na(y))))
})
na_counts
```

```{r}
full_matrix <- df %>% 
    mutate(income_level = if_else(
        income_level == ">50K", 
        1L, 
        0L)) %>% 
    mutate_if(is.character, as.factor) %>% 
    data.matrix()

sparse_matrix <- sparse.model.matrix(
    income_level ~ ., 
    data = df %>% 
        mutate(income_level = if_else(
            income_level == ">50K", 
            1L, 
            0L)))
    
head(sparse_matrix)
output_vector <- as.integer(df[, "income_level"] == ">50K")
head(output_vector, n = 50)
train_indxs <- createDataPartition(df$income_level, 
                                   times = 1, 
                                   p = 0.8, 
                                   list = FALSE)

train <- full_matrix
```

```{r}

df %>% 
    mutate(income_level = if_else(
        as.integer(income_level) == 2 ~ 1L, 
        TRUE                          ~ 0L
    ))
# DT[as.integer(income_level) == 2]
unique(df$income_level)
numeric_df <- data.matrix(df)
unique(numeric_df[, "income_level"])
sparse_matrix <- sparse.model.matrix(DT)

data.frame(a = df$income_level[1:10], 
           b = numeric_df[1:10, "income_level"])

```


```{r}

```

[1]: https://archive.ics.uci.edu/ml/datasets/adult